\documentclass[12pt]{article}
\renewcommand{\baselinestretch}{1.2}
\begin{document}

The following is a list of suggested problems to work on.  Obviously
there won't be this many on the exam!  But since I never gave you an
explicit list of problems to look at, I thought I'd do that now.  I
didn't include any of the  ``review'' problems at the end of the
chapters.  So if you want problems after working through these, you
can do those also.  Enjoy!

Problems from handouts:
\begin{itemize}
\item chapter 1-2: These questions are ``understanding'' questions.  I
will probably write a question similar to one of these (where I'll
start with a newspaper article).  You should feel confortable with all
of these questions.
\item chapter 16: 
  \begin{itemize}
  \item page 277: 4,  8, 9.
  \item page 280: 4, 6
  \item page 284: 2, 5, 8, 10
  \end{itemize}
\item chapter 17: 
  \begin{itemize}
  \item page 290: 2, 4,
  \item page 293: 3, 6
  \item page 296: 1, 8
  \item page 299: 4
  \item page 303: 4, 9
  \end{itemize}
\item chapter 18:
  \begin{itemize}
  \item page 313: 4, 5
  \item page 318: 4
  \item page 324: 5, 9
  \end{itemize}
\item chapter 26: 
\begin{itemize}
\item page 478: 3, 5
\item page 479: 3, 5
\item page 482: 3, 7
\item page 484: 1, 5
\item page 487: 1, 10
\item page page 495: 6
\end{itemize}
\item chapter 27:
  \begin{itemize}
  \item page 505: 5
  \item page 508: 2, 3
  \item page 513: 2
  \item page 516: 4, 5, 7
  \end{itemize}
\end{itemize}
MLE problems:
\begin{itemize}
\item Let $X_1$, $X_2$, $\ldots$, $X_n$ be a sample from a population
with a Pareto density:
$$f(x,\theta) = \theta x^{-(\theta+1)}$$
Find the MLE of $\theta$.  (Note: $x \ge 1$.)
\item Let $X_1$, $X_2$, $\ldots$, $X_n$ be a sample from a population
with a Beta density:
$$f(x,\theta) = \sqrt\theta x^{(\sqrt\theta-1)}$$
Find the MLE of $\theta$.  (Note: $0 \le x \le 1$.)
\item Let $X_1$, $X_2$, $\ldots$, $X_n$ be a sample from a population
with a Rayleigh density:
$$f(x,\theta) = (x/\theta^2)e^{-x^2/2\theta^2}$$
Find the MLE of $\theta$.  (Note: $0 \le x$.)
\end{itemize}


\end{document}
