\documentclass[12pt]{extarticle}


\begin{document}

{\bf Interval Questionnaire.}

\begin{enumerate}
\item Distance in air miles between Chicago and Paris.

\item Number of lakes in the state of Michigan.

\item Average life length (in years) of a giraffe.

\item Length of the Nile River.

\item Number of countries that are members of OPEC.

\item Life expectancy of a baby boy born in 1979 in the U.S.

\item Number of books in the Old Testament.

\item Diameter of the moon in miles.

\item The year that Richard Nixon was born.
\item
 Number of pennies needed to weigh one pound.
\item
 Number of signatories to the Declaration of Independence.
\item
 Number of bound volumes in the libraries of the University of
Chicago in 1986.
\item
 Gestation period (in days) of a female Asian elephant.
\item
 The cost of a full-membership seat on the Chicago Board of Trade
at the start of 1984.
\item
Deepest point in the oceans (in feet).
\item
The number of stores in Chicago listed as toy stores in the 1986
yellow pages. 
\item
Year in which the Book of Mormon was first published.
\item
Population of the State of Hawaii (1980 census).
\item
How much money was spent on tranquilizers in the U.S. during
1980. 
\item
Height of the world's highest waterfall.  
\end{enumerate}


  This quiz was created at the University of Chicago by J. Klayman.


\newpage

(1) 4337 miles, (2) 11,000
  lakes, (3) 10 years, (4) 4145 miles, (5) 13, (6) 72 years, (7) 39
  books, (8) 2160 miles, (9) 1913, (10) A gross (i.e. 144) (11) 56,
  (12) 6-7 million volumes, (13) 645 days (14) \$300,000, (15) 36,198
  feet, (16) 31 toy stores (17) 1830, (18) 915,000, (19) \$2.3 Billion
  and (20) 3212 ft (in Venezuela).

  In case your math isn't too good, 95\% of 20 questions is 19.  So,
  you should have gotten about 1 question wrong.  If your ``normal''
  you got more than 1/2 of them wrong.  In other words, your coverage
  frequency was less than 50\% instead of being 95\%. 

  It is interesting that if you ask people after they fill in the
  questionnaire how many they think they got correct, they guess closer
  to 50\% than 95\%.  In spite of what the instructions told them!


\end{document}
